---
title: "Regression Project"
output: word_document
date: "2023-12-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(MASS)
library(psych)
library(tinytex)
library(knitr)
library(olsrr)
```

```{r}
my_data<-read.csv("insurance.csv",header=TRUE)
attach(my_data)
head(my_data,2)
```

```{r}
plot(my_data)
pairs.panels(my_data)
names(my_data)
```

- Running the regression Analysis
- Modifying the model since R doesn't recognize categorical and integer variables 
```{r}
gender <- as.factor(sex)
smokes <- as.factor(smoker)
place <- as.factor(region)
my_data<-cbind(my_data,gender,smokes,place)
children <- as.numeric(children)

m1<-lm(expenses~age+bmi+children+sex+smoker+region)
summary(m1)
par(mfrow=c(2,2))
plot(m1)
```
Residual standard error is very high --> modify the model

```{r}
xx<-boxcox(m1)
xx$x[which.max(xx$y)]
```

TRANSFORMATION (POWER 0.1414)
```{r}
m2<-lm(expenses^0.1414~age+bmi+children+gender+smokes+place)
summary(m2)
par(mfrow=c(2,2))
plot(m2)
```


Plotting the Interaction plots to identify the interaction variables
```{r}
interaction.plot(gender,smokes,expenses)
interaction.plot(gender,place,expenses)
interaction.plot(smokes,place,expenses)
```

Adding the interaction terms
```{r}
m3<-lm(expenses^0.1414~age+bmi+children+ gender*smokes+gender*place+ smokes*place)
summary(m3)
par(mfrow=c(2,2))
plot(m3)
```


Calculating hadi's measure and cooks distance to identify the influential points

AFTER THE TRANSFORMATION AND INTERACTION TERMS
```{r}
row.names(my_data)=1:nrow(my_data)

hadi<-ols_hadi(m3)
cook<-cooks.distance(m3)
hat<-hatvalues(m3)

ols_plot_hadi(m3)

par(mfrow=c(2,1))

plot(cook,main="cook's distance")

plot(hat,main="hatvalues")

#identify(row.names(my_data),hadi$hadi,my_data,labels = row.names(my_data)) #this command to be implemented in the console
```
We Identified that the there are 4 influential points from hadi and so we need to identify the corresponding observation numbers
```{r}
print("influentials from Hadi's measure")
tail(sort(hadi$hadi),4)

print("influentials from cook's distance")
tail(sort(cook),4)
```

Removing the influential points identified from hadi's measure
```{r}
my_data1<-my_data[-c(1028,431,220,517),]
m4<-lm(expenses^0.1414~age+bmi+children+gender*smokes+gender*place+ smokes*place,data=my_data1)
summary(m4)
par(mfrow=c(2,2))
plot(m4)
```
```{r}
row.names(my_data)=1:nrow(my_data)

hadi<-ols_hadi(m4)
cook<-cooks.distance(m4)
hat<-hatvalues(m4)

ols_plot_hadi(m4)

par(mfrow=c(2,1))

plot(cook,main="cook's distance")

plot(hat,main="hatvalues")
```

Comparing the 2 models
```{r}
anova(m2,m3)
```
P-value<0.05
Reject H0 and accept H1 = Model 2 is better



```{r}
ols_coll_diag(m4)
ols_step_backward_p(m4) #identifying model 4
par(mfrow=c(2,2))
plot(m4, main = "Before the step function")
m5 <- step(m4,direction="backward",test="F") #identifying model 5
plot(m5, main = "After the step function")
summary(m5)
```


VIF < 10 --> no collinearity
Condition Index for 2 sets of variables (14 & 15) was > 10
No variables with a value > 0.7
Step function removed the sets of variables with condition index > 10 (from the interaction term)


```{r}
anova(m5,m4)
```
H0: Model 5 = Reduced Model     VS          H1: Model 4 = Full Model
P-value > 0.05
Fail to Reject H0; Model 1 is significant; after removing (gender:place) interaction term the model is more adequate = worth losing the degrees of freedom



